---
title: 'Appendix: Interrupted Time Series When the Outcome Variable is a Moving Average'
output:
  html_notebook:
    df_print: paged
---

The data provided by the Department of Health for many outcome variables are 12-month trailing moving averages of monthly values. DOH is unable to provide the underlying monthly values. The moving-average outcomes data require special treatment when conducting interrupted time series (ITS) analysis to ensure that we properly estimate the impact of DSRIP. This appendix explains the method we use and why.

NOTE: This appendix is fully reproducible by running the R program file "appendix_its_ma.rmd".


# Issues created by moving average data
The moving-average outcomes data create two main issues:

1. **Lagged and muted effects**:  Effects from the DSRIP policy intervention will appear only gradually in the moving-average values of the outcome. These effects will be misestimated and likely underestimated if we use a typical ITS model without adjustment.
2. **Autocorrelated errors**: Simple models of the moving-average outcomes will have errors that are correlated over time.

Both issues can be addressed.

# A way to address these issues
If we had the "true" data (the underlying monthly values, rather than the 12-month moving averages), we could estimate a simple ITS model. Assuming (for now) a linear trend and no seasonality, we might estimate the following simple model:

y = b0 + b1 * time + b2 * policy + b3 * policy * time + e

where y is the actual monthly variable, time is a trend variable, and policy is an indicator variable that is 1 beginning when DSRIP went into effect and 0 in every prior month. The coefficient b2 allows for a change in intercept (for example, an increase in the outcome after DSRIP goes into effect) and b3 allows for a change in slope after DSRIP goes into effect.

However,if we do not have actual monthly values for y but only have 12-month trailing moving-average values (y_ma), this model is inappropriate. 

For reasons we give below, we instead estimate a model in which the dependent variable is a moving average and the independent variables are moving averages of the independent variables above.


# Explanation of this approach
Here we show the proper model. Begin by adding a time subscript for all variables in the true model so that the value of y for a particular time is:

y[t] = b0 + b1 * time[t] + b2 * policy[t] + b3 * policy[t] * time[t] + e[t]

The value for the moving average of the outcome at time t is y_ma[t]. This is calculated as:

y_ma[t] =

          (y[t] +

           y[t-1] +
           
           y[t-2] +
           
           ...
           
           y[t-11]) / 12
           
This can be rewritten as the average of the righthand side of the true model above, or:

y_ma[t] = 

          (b0 + b1 * time[t]    + b2 * policy[t]    + b3 * policy[t]    * time[t]    + e[t] +
          
           b0 + b1 * time[t-1]  + b2 * policy[t-1]  + b3 * policy[t-1]  * time[t-1]  + e[t-1] +
           
           b0 + b1 * time[t-2]  + b2 * policy[t-2]  + b3 * policy[t-2]  * time[t-2]  + e[t-2] +
           
           ....
           
           b0 + b1 * time[t-11] + b2 * policy[t-11] + b3 * policy[t-11] * time[t-11] + e[t-11]) / 12
           

Rearranging terms and dividing each term by 12 gives:

y_ma[t] = 

         (12 * b0 / 12) +
         
               b1 * (time[t] + time[t-1] + time[t-2] + ... + time[t-11]) /12 +
               
               b2 * (policy[t] + policy[t-1] + policy[t-2] + ... + policy[t-11]) /12 +
               
               b3 * (policy[t] * time[t] + policy[t-1] * time[t-1] + policy[t-2] * time[t-2] + ... + policy[t-11] * time[t-11]) /12 +
               
                    (e[t] + e[t-1] + e[t-2] + ... + e[t-11]) /12

           
Which simplifies to:

y_ma[t] = b0 + b1 * ma12(time[t]) + b2 * ma12(policy[t]) + b3 * ma12(policy[t] * time[t]) + ma12(e[t])

where ma12(x[t]) is the 12-month trailing moving average of the variable x in time period t.


Thus, the coefficients in a model with the moving-average of the outcome as the dependent variable and moving averages of the independent variables plus an autocorrelated error term, are the same as the coefficients in the true model.

**This is nice from the perspective of interpretation and communication. We can present the same kinds of coefficients and standard errors as are in traditional ITS presentations, and as are in the interim evaluation report now, and they can be interpreted easily. And we don't need to present the autocorrelation term if we don't want to. (I wouldn't.) We can just make sure we say that we did it. We can have a footnote that says available upon request, or something like that.**

There are two more things we can do to make interpretation easy for our readers:

1. Define the time or trend variable so that the final pre-intervention month is month number 0, and the first intervention month is number 1. Thus, if we have 13 months of pre-intervention data and 24 months of intervention data (37 observations), the pre-intervention period will be numbered from -12 to 0 (13 points), and the intervention period will be numbered from 1 to 24. This makes interpretation of the coefficients really easy as you will see below.

2. Create graphs that show the implications of just the intervention coefficients (policy + policy * time in the graphs below). This shows the intervention effect nicely.

The analysis below with illustrative data shows that this works well.

Thus, in this project we estimate this model.


# Illustration with constructed data of the appropriateness of the approach
In this section we create an illustrative dataset that has actual monthly values of an outcome variable for which we know the true impact of a policy intervention. We create a moving average of the outcome variable. 

We then estimate ITS models using a simple method plus two versions of the method we describe above, and show that the method we use produces results that are much closer to the true impact of the policy than those from the simple model.

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library("magrittr")
library("plyr") # needed for ldply; must be loaded BEFORE dplyr
library("tidyverse")
options(tibble.print_max = 60, tibble.print_min = 60) # if more than 60 rows, print 60 - enough for states
# ggplot2 tibble tidyr readr purrr dplyr stringr forcats

library("scales")
library("hms") # hms, for times
library("lubridate") # lubridate, for date/times
library("vctrs")

library("grDevices")
library("knitr")
library("kableExtra")

library("zoo") # for rollapply

library("forecast") # for auto.arima

```


```{r functions, message=FALSE, warning=FALSE, include=FALSE}
ma <- function(x, period) {
  # create trailing moving average of x of length period
  zoo::rollapply(x, period, function(x) mean(x, na.rm=TRUE), fill=NA, align="right")
}

```


## Create a small dataset similar to data used in the evaluation
We create an illustrative data set with 60 monthly observations (5 years), 36 of which are BEFORE DSRIP goes into effect and 24 of which are after DSRIP goes into effect

The set data has the following variables:

* y: the monthly value of the outcome variable - the true value, not a moving average
* time: we define this so that it is 1 in the first year in which DSRIP is in effect. Thus, the 
  value of time 
* policy is an indicator of when the policy change went into effect. It is 1 for time


In our example data we KNOW the true effect of the policy, which follows an ITS model

y = b0 + b1*time + b3*policy + b4*policy*time


```{r define_impact}
# baseline model for outcome variable
b0 <- 100 # intercept
b1 <- -1 # slope

# policy impacts
b2 <- 2 # policy intercept shift
b3 <- 1.5 # policy slope shift

errSD <- 2.5 # SD of random error

set.seed(3)

```


```{r create_data, message=FALSE, warning=FALSE, include=FALSE}
params <- c(b0, b1, b2, b3)

dfraw <- tibble(time=-35:24,
             policy=ifelse(time > 0, 1, 0),
             shift_effect=b2 * policy,
             slope_effect=b3 * policy * time,
             y_true = b0 + b1*time + shift_effect + slope_effect, # the true unknown value
             err=rnorm(n=length(time), sd=errSD), # note that errors are homoscedatic by design - not likely in real data
             y = y_true + err) # the observed value

# quick look to make sure the data seem useful
dfraw %>% 
  gather(variable, value, starts_with("y")) %>%
  ggplot(aes(time, value, colour=variable)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept=0.5, linetype="dashed") +
  geom_hline(yintercept = 100, linetype="dashed")

# add moving averages
dfraw <- dfraw %>%
  mutate(y_ma=ma(y, 12),
         policy_ma=ma(policy, 12),
         time_ma=ma(time, 12),
         joint_ma=ma(time * policy, 12),
         err_ma=ma(err, 12))

```


The plot below shows y and its moving average.

```{r yyma_plot, echo=FALSE, message=FALSE, warning=FALSE}
# plot y and its moving average
p <- dfraw %>% 
  # filter(time %in% -6:12) %>% # zoom in on the phase-in period
  gather(variable, value, y, y_ma) %>%
  mutate(variable=factor(variable, 
                         levels=c("y", "y_ma"), 
                         labels=c("y true", "y moving\n average"))) %>%
  ggplot(aes(time, value, colour=variable)) +
  geom_line() +
  geom_point() +
  geom_vline(xintercept=0.5, linetype="dashed") +
  # geom_hline(yintercept = 100, linetype="dashed") +
  scale_x_continuous(name="Month (1=first month of policy change)", breaks=seq(-36, 48, 3)) +
  scale_y_continuous(name="Outcome", breaks=seq(0, 200, 5)) +
  annotate("text", x=12, y=120, label="Moving average\nshows less improvement\nthan true value") +
  theme_bw() +
  theme(legend.title = element_blank()) +
  ggtitle("True outcome variable and its 12-month moving average")
p

```


```{r eval=FALSE, include=FALSE}
Acf(dfraw$y)
Acf(dfraw$y_ma)

Pacf(dfraw$y)
Pacf(dfraw$y_ma)

```



## ITS model estimated on the true data
```{r mod_true, echo=FALSE, message=FALSE, warning=FALSE}
# first, estimate the true model to see how much impact err has
m_mod0 <- lm(y ~ time + policy + I(policy*time), data=dfraw)
summary(m_mod0)
# show parameters
params 

# Acf(residuals(m_mod0))

```


## Estimate ITS models using the moving-average data
```{r mod_data, echo=FALSE, message=FALSE, warning=FALSE}
# create data frame with just the observations we need
dfmod <- dfraw %>% 
  filter(!is.na(y_ma))

```


### The simple model
```{r simple, echo=FALSE, message=FALSE, warning=FALSE}
m_mod1 <- lm(y_ma ~ time + policy + I(policy*time), data=dfmod)
summary(m_mod1)
params
# Acf(residuals(m_mod1))

# plot the policy effects over 2 years

effects <- tibble(time=1:24) %>%
  mutate(emod0=coef(m_mod0)["policy"] + coef(m_mod0)["I(policy * time)"] * time,
         emod1=coef(m_mod1)["policy"] + coef(m_mod1)["I(policy * time)"] * time)

levs <- c("emod0", "emod1")
labs <- c("Model using actual monthly data",
              "Model using moving average (MA) only for dependent variable (interim report)")

p <- effects %>%
  gather(variable, value, -time) %>%
  mutate(variable=factor(variable,
                         levels=levs,
                         labels=str_wrap(labs, 25))) %>%
  ggplot(aes(time, value, colour=variable)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=0) +
  scale_x_continuous(name="Month (1=first month of policy change)", breaks=seq(-36, 48, 3)) +
  scale_y_continuous(name="Outcome", breaks=seq(0, 200, 5)) +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.text = element_text(margin = margin(t = 5, b = 5, unit = "pt"))) +
  ggtitle("Policy effects estimated from simple model, first 2 years of effect")
p

```


### Model with preferred independent variables
```{r mod_goodrhs, echo=FALSE, message=FALSE, warning=FALSE}
# run the proper model without autoregressive term
m_mod2 <- lm(y_ma ~ time_ma + policy_ma + joint_ma, data=dfmod)
summary(m_mod2)
# Acf(residuals(m_mod2))
# Pacf(residuals(m_mod2)) # ** the high correlation at lag 1 suggests AR(1) **

```

### Preferred model: preferred independent variables AND autocorrelation correction
Finally, correct for autocorrelation.

```{r mod_ar, echo=FALSE, warning=FALSE, message=FALSE}
# what if we actually knew the error term?
m_mod2a <- lm(y_ma ~ time_ma + policy_ma + joint_ma + err_ma, data=dfmod)
summary(m_mod2a)
params

# but in reality we won't know it, so we add an AR term to the model
xvars <- as.matrix(dfmod[, c("time_ma", "policy_ma", "joint_ma")])

m_mod3 <- arima(dfmod$y_ma, xreg=xvars, order=c(1,0,0))
m_mod3
# Acf(residuals(m_mod3))
# 
m_mod4 <- auto.arima(dfmod$y_ma, xreg=xvars)
m_mod4
 #Acf(residuals(m_mod4))
 #Pacf(residuals(m_mod4))


m_mod5 <- arima(dfmod$y_ma, xreg=xvars, order=c(0,0,11), fixed = c(rep(1, 11), NA, NA, NA, NA ))
m_mod5

Acf(residuals(m_mod5))
Pacf(residuals(m_mod5))


effects <- tibble(time=1:24) %>%
  mutate(etrue=b2 + b3 * time,
         emod0=coef(m_mod0)["policy"] + coef(m_mod0)["I(policy * time)"] * time,
         emod1=coef(m_mod1)["policy"] + coef(m_mod1)["I(policy * time)"] * time,
         emod2=coef(m_mod2)["policy_ma"] + coef(m_mod2)["joint_ma"] * time,
         emod3=coef(m_mod3)["policy_ma"] + coef(m_mod3)["joint_ma"] * time,
         emod4=coef(m_mod4)["policy_ma"] + coef(m_mod4)["joint_ma"] * time,
         emod5=coef(m_mod5)["policy_ma"] + coef(m_mod5)["joint_ma"] * time)

models <- c("etrue", "emod0", "emod1", "emod2", "emod3", "emod4", "emod5")
modnames <- c("Model with true coefficients and no error",
              "Model using actual monthly data",
              "Model using moving average (MA) only for dependent variable (interim report)",
              "Model using MA for depvar and for independent variables",
              "Model using MA for depvar and for independent variables, plus AR term (preferred)",
              "Model using MA for depvar and for independent variables, plus ARMA",
              "Model using MA for depvar and for independent variables, restricted MA")
names(modnames) <- models

```


## Month-by-month estimated impacts, alternative models
```{r echo=FALSE, warning=FALSE, message=FALSE}
vnames <- c("time", models)
colnames <- c("time", modnames)
names(colnames) <- vnames
# colnames["emod3"]
# Note: ok to put html breaks in colnames -- e.g., "very<br/>long name"

effects  %>%
  select(time, emod0, emod1, emod2, emod3, emod5) %>%
  kable(format="html", col.names=colnames[names(.)], escape=FALSE, digits=2)

```


```{r echo=FALSE, warning=FALSE, message=FALSE, fig.height=6, fig.width=10}
levs <- c("emod0", "emod1", "emod2", "emod3", "emod5")
labs <- modnames[levs]

p <- effects %>% 
  select(time, emod0, emod1, emod2, emod3, emod5) %>%
  gather(variable, value, -time) %>%
  mutate(variable=factor(variable,
                         levels=levs,
                         labels=str_wrap(labs, 25))) %>%
  ggplot(aes(time, value, colour=variable)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=0) +
  scale_x_continuous(name="Month (1=first month of policy change)", breaks=seq(-36, 48, 3)) +
  scale_y_continuous(name="Outcome", breaks=seq(0, 200, 5)) +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.text = element_text(margin = margin(t = 5, b = 5, unit = "pt"))) +
  ggtitle("Estimated impact of policy under different modeling approaches")
p


p2 <- effects %>% 
  select(time, emod0, emod5) %>%
  gather(variable, value, -time) %>%
  mutate(variable=factor(variable,
                         levels=levs,
                         labels=str_wrap(labs, 25))) %>%
  ggplot(aes(time, value, colour=variable)) +
  geom_line() +
  geom_point() +
  geom_hline(yintercept=0) +
  scale_x_continuous(name="Month (1=first month of policy change)", breaks=seq(-36, 48, 3)) +
  scale_y_continuous(name="Outcome", breaks=seq(0, 200, 5)) +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.text = element_text(margin = margin(t = 5, b = 5, unit = "pt"))) +
  ggtitle("Estimated impact of policy under different modeling approaches")
p2


```



# Policy-effect coefficients, and effects calculated for months 1 and 24
```{r tab, echo=FALSE, message=FALSE, warning=FALSE}
evals <- tibble(type="Calculation with true coefficients (no error)", intercept=b2, slope=b3) %>%
  bind_rows(tibble(type=modnames["emod0"], intercept=coef(m_mod0)["policy"], slope=coef(m_mod0)["I(policy * time)"]),
            tibble(type=modnames["emod1"], intercept=coef(m_mod1)["policy"], slope=coef(m_mod1)["I(policy * time)"]),
            tibble(type=modnames["emod2"], intercept=coef(m_mod2)["policy_ma"], slope=coef(m_mod2)["joint_ma"]),
            tibble(type=modnames["emod3"], intercept=coef(m_mod3)["policy_ma"], slope=coef(m_mod3)["joint_ma"]),
            tibble(type=modnames["emod4"], intercept=coef(m_mod4)["policy_ma"], slope=coef(m_mod4)["joint_ma"]),
            tibble(type=modnames["emod5"], intercept=coef(m_mod5)["policy_ma"], slope=coef(m_mod5)["joint_ma"])) %>%
  mutate(effect_month1=intercept + slope, 
         effect_month24=intercept + slope * 24)

evals %>%
  kable(digits = 2)

```



